---
title: "Company Growth"
format: pdf
editor: visual
---

```{r setup}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(vroom)
library(zoo)
library(forecast)
library(splines)

data <- vroom('CompanyGrowth.csv')
data <- data %>%
  mutate(date = as.Date(as.yearqtr(paste(Year, Qrtr, sep = "-"))))
```

## EDA

```{r}
# growth stays about the same throughout
ggplot(data=data, aes(x=date, y=PctGrowth)) +
  geom_point()
# Change in Income
ggplot(data=data, aes(x=Income, y=PctGrowth)) +
  geom_point()
# Change in Production
ggplot(data=data, aes(x=Production, y=PctGrowth)) +
  geom_point()
# Change in Savings
ggplot(data=data, aes(x=Savings, y=PctGrowth)) +
  geom_point()
# Change in Unemployment
ggplot(data=data, aes(x=Unemployment, y=PctGrowth)) +
  geom_point()
```

1.  Describe the model you are currently using, how it performs for predictions and what the 2025 projections are under this model. Be sure to describe how you obtained the 2025 predictions. Company Growth Forecasting Problem Background Growth = β0 + β1Income + β2Production + β3Savings + β4Unemployment + ϵi ϵi iid ∼ N (0, σ2)

    ```{r}
    # basic linear model
    growth.lm <- lm(PctGrowth ~ Income + Production + Savings + Unemployment, data=data)

    # lm models for each predictor
    new_data <- data.frame(Year = 2025,
                           Qrtr = c(1,2,3,4))
    income.lm <- lm(Income~Year+Qrtr, data=data)
    income_lm_preds <- predict(income.lm, newdata = new_data)
    production.lm <- lm(Production~Year+Qrtr, data=data)
    production_lm_preds <- predict(production.lm, newdata = new_data)
    savings.lm <- lm(Savings~Year+Qrtr, data=data)
    savings_lm_preds <- predict(savings.lm, newdata = new_data)
    unemployment.lm <- lm(Unemployment~Year+Qrtr, data=data)
    unemployment_lm_preds <- predict(unemployment.lm, newdata = new_data)

    # predict growth
    growth_new_data <- data.frame(Income = income_lm_preds,
                                  Production = production_lm_preds,
                                  Savings = savings_lm_preds,
                                  Unemployment = unemployment_lm_preds)
    predict(growth.lm, growth_new_data)

    ## how well does our model predict
    # r^2
    summary(growth.lm)$r.squared
    # rmse
    rmse <- sqrt(mean(residuals(growth.lm)^2))
    rmse
    ```

2.  Describe at least 2 alternative models that could be used to do the projections and emphasize how they differ from the current method.

    ```{r}
    # new linear model accounting for date features and maybe with spline and what not
    data_date_features <- data %>%
      mutate(Year = year(date),
             Month = month(date, label = TRUE),
             DOW = wday(date, label = TRUE),
             Numeric_date = decimal_date(date))

    spline_model <- lm(PctGrowth ~ bs(Numeric_date, knots = c(1975, 1990, 2005, 2020), degree = 3) +
    ., data = data_date_features)

    data_date_features %>%
    ggplot(aes(x = date, y = PctGrowth)) +
    geom_point() +
    geom_line(aes(y = spline_model$fitted.values), col = 'red')

    future_dates <- seq(from = as.yearqtr("2025 Q1"), to = as.yearqtr("2025 Q4"), by = 1/4)
    future_dates <- as.Date(as.yearqtr(future_dates), frac = 0)
    future_features <- tibble(date = future_dates) %>%
      mutate(Year = year(date),
        Month = month(date, label = TRUE),
        DOW = wday(date, label = TRUE),
        Numeric_date = decimal_date(date),
        Qrtr = quarter(date, with_year = FALSE, fiscal_start = 1),
        Income = income_lm_preds,
        Production = production_lm_preds,
        Savings = savings_lm_preds,
        Unemployment = unemployment_lm_preds)

    predict(spline_model, newdata = future_features)

    summary(spline_model)$r.squared

    rmse_spline <- sqrt(mean(residuals(spline_model)^2))
    rmse_spline
    # facebook prophet (maybe or maybe not this one because it won't account for other variables)

    ```

    ```{r}
    # ARIMA
    ## Define response as time series object with a frequency S
    ts_data <- ts(data=data$PctGrowth, frequency=4)
    X_noInt <- model.matrix(growth.lm)[,-1]
    arima_model <- auto.arima(y=ts_data,xreg=X_noInt)
    summary(arima_model)
    coef(arima_model)
    ```

3.  Compare these 2 new methods to the current method in terms of ability to predict and the 2025 projections. Be sure to describe how you obtained the 2025 predictions. (predict for all variables, not just the response)

    ```{r}

    ```

4.  The board is particularly interested to know how Income, Production, Savings and Unemployment relate to company growth. Use your preferred method (among the 3 or more you consider here) to illustrate this relationship.
